---
title: "TIL Invasion Descriptive Statistics"
date: "`r Sys.Date()`"
output: 
   html_document:
      toc: yes
      toc_depth: 4
   pdf_document:
      toc: yes
      toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/data")
approxHeight = params$nVars
knitr::opts_chunk$set(collapse = T, warning = F, fig.width=10, fig.height=approxHeight, echo = F)
library(readr)
library(plyr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(survival)
library(survminer)
```

\newpage

```{r load, include = F}
df =  as.data.frame(readr::read_csv(file = params$csvPath))
toFacet = colnames(df)[which(!colnames(df) %in% c("scaled_PP","TIL_Class",params$survCensor, params$survTime))]
```

# Descrpitive Statistics

## Overall Distribution

```{r, fig.height = 4}
ggpubr::gghistogram(df, x = "scaled_PP", fill = "TIL_Class", binwidth = .1)
```

In the analyzed dataset, you have provided **`r nrow(df)` samples** to be analyzed and **`r length(toFacet)` columns** to subdivide the samples. The pipeline calls Low TIL vs High TIL around the **mean** of the dataset (will print mean here). This results in

* `r sum(df$TIL_Class == "Low")` Low invasion samples (`r signif((sum(df$TIL_Class == "Low")/nrow(df))*100,digits = 4)`%) and
* `r sum(df$TIL_Class == "High")` High invasion samples (`r signif((sum(df$TIL_Class == "High")/nrow(df))*100,digits = 4)`%). 

Values for invasion were not computed for **`r sum(is.na(df$scaled_PP))` samples**. Below we will compare invasion characteristics across your provided additional features (all additional columns beyond survival, scaled_PP, and TIL_Class) and, if requested, perform some clinical correlations with your outcome information. 

\newpage

## Faceted by variables of interest

### Class of Invasion

Here we test if invasion class is independent of your variable of interest using a chi-squared test

```{r}
plots = list()
for(i in toFacet){
   tmp = df[,c(i, "TIL_Class")]
   
   chi = table(tmp[,i],tmp$TIL_Class)
   chi = chisq.test(chi)
   # 
   tmp = tmp %>% plyr::count()
   plots[[i]] = ggpubr::ggbarplot(tmp, x = i, y = "freq",
                                  fill = "TIL_Class") +
      labs(y = "Num of WSI",
           caption = paste0("Chisq pval: ", signif(chi$p.value,3), "\n",
                            "Chi-stat: ", signif(chi$statistic,6)))
}
x = ggarrange(plotlist = plots, ncol = length(toFacet), common.legend = T)
x
```

### Continuous Invasion

For continuous invasion, we use a non-parametric Wilcoxon Rank-Sum if your provided feature has 2 levels. If there are 3+ levels, ANOVA is used.

```{r}
plots = list()
for(i in toFacet){
   if(length(levels(as.factor(df[,i]))) > 2){
      plots[[i]] = ggpubr::ggviolin(df, x = i,
                                    y = "scaled_PP",
                                    fill = i, trim = T) +
         stat_compare_means(method = "anova",
                            label.x.npc = "center")
   } else {
      plots[[i]] = ggpubr::ggviolin(df, x = i,
                                    y = "scaled_PP",
                                    fill = i, trim = T) +
         stat_compare_means(method = "wilcox",
                            label.x.npc = "center")
   }
}
ggarrange(plotlist = plots, ncol = length(toFacet))
```

\newpage

# Survival

This section assumes you provided the column names in your csv for "Time to Event" and "Is sample censored (0 = yes, 1 = no). Please see the analytic file help for details if this is not clear.

## Kalpan-Meier

For the Kaplan-Meier section, TIL Class is used to categorize samples and the log-rank test is used to compute p-values.

### Univariate

```{r, fig.height = 6, eval = params$includeSurv}
tmp.df = df
colnames(tmp.df)[which(colnames(tmp.df)==params$survTime)] = "survTime"
colnames(tmp.df)[which(colnames(tmp.df)==params$survCensor)] = "survCensor"
tmp.df$SurvObj <- with(tmp.df, Surv(time = survTime,
                                    event =survCensor))
fit <- survfit(SurvObj ~ TIL_Class,
               data = tmp.df,
               type = "kaplan-meier")

km = ggsurvplot(fit,
                data = tmp.df,
                #palette = c("#00BFC4","#F8766D"),
                risk.table = T,
                pval = TRUE,
                censor = TRUE,
                surv.mean.line = "hv",
                legend.title = "TIL Infiltration") + ggtitle("Survival by TIL Class")

km
```

\newpage

### By Vars of Interest

```{r, fig.height = 8, fig.width = 12, eval = params$includeSurv}
plots = list()
for(i in toFacet){
   km = ggsurvplot_facet(fit,
                         data = tmp.df,
                         risk.table = T,
                         facet.by = i,
                         pval = TRUE,
                         censor = TRUE,
                         surv.mean.line = "hv",
                         legend.title = "TIL Infiltration",
                         xlab = "Time") + ggtitle(paste0("Survival by TIL Class + ", i))
   plots[[i]] = km
}

ggarrange(plotlist = plots, common.legend = T)
```

\newpage

## Cox Regression

For the Cox regression section, both TIL Class (categorical) and scaled_PP (continuous) are used to categorize samples. While KM plots are limited to bivariate, this section will run all analyses in both bivariable and in a larger model using all included variables.

### Univariate Categorical

```{r, fig.height = 3, fig.width = 8, eval = params$includeSurv}
tmp.df$TIL_Class = factor(tmp.df$TIL_Class, levels = c("Low","High"))
fit.coxph <- coxph(Surv(time = tmp.df$survTime,
                        event = tmp.df$survCensor,
                        type = "right") ~ TIL_Class, data = tmp.df)
x = ggforest(fit.coxph, data = tmp.df, main = "TIL Class")
x
```

### Univariate Continuous

```{r, fig.height = 2, fig.width = 8, eval = params$includeSurv}
fit.coxph <- coxph(Surv(time = tmp.df$survTime,
                        event = tmp.df$survCensor,
                        type = "right") ~ scaled_PP, data = tmp.df)
x = ggforest(fit.coxph, data = tmp.df, main = "Continuous Invasion (scaled by SD)")
x
```

### By Vars of Interest (one at a time)

```{r, fig.width = 8, eval = params$includeSurv}
plots = list()
for(i in toFacet){
   newTmp = tmp.df
   #colnames(newTmp)[which(colnames(newTmp) == i)] = "Var"
   fit.coxph <- SurvObj ~ .
   toUse = paste0("scaled_PP+",i)
   fit.coxph = reformulate(toUse, fit.coxph[[2]])
   fit.coxph = coxph(fit.coxph, data = newTmp)
   plots[[i]] = ggforest(fit.coxph, data = newTmp, main = paste0("Bivariate with ", i))
}
ggarrange(plotlist = plots, nrow = length(toFacet))
```

### Multivariable Cox (all features)

```{r, fig.width = 8, eval = params$includeSurv}
newTmp = tmp.df[,c("scaled_PP", "SurvObj", toFacet)]
fit.coxph <- coxph(SurvObj ~ ., data = newTmp)
x = ggforest(fit.coxph, data = newTmp, main = "All Var")
x
```
